Supplemental materials for \textit{Category-based and location-based volitional covert attention affect memory at different timescales}

### 2023-03-27 paper/main.tex
minor wording tweaks, typo and formatting fixes

Our ongoing subjective experiences, and our memories of those experiences, are
shaped by our prior experiences, goals, and situational understanding. These
factors shape how we allocate our attentional resources over different aspects
of our ongoing experiences. These attentional shifts may happen overtly (e.g.,
when we change where we are looking) or covertly (e.g., without any explicit
physical manifestation). Additionally, we may attend to what is happening at a
specific spatial location (e.g., because we think something important is
happening there) or we may attend to particular features irrespective of their
locations (e.g., when we search for a friend's face in a crowd versus a
location on a map). We ran a covert attention experiment with two conditions
that differed in how long they asked participants to maintain the focus of the
categories and locations they were attending. Later, the participants performed
a recognition memory task for attended, unattended, and novel stimuli.
Participants were able to shift the location of their covert attentional focus
more rapidly than they were able to shift their focus of covert attention to
stimulus categories, and the effects of location-based attention on memory were
longer-lasting than the effects of category-based attention.

@@

We ran a total of 53 participants in across two experimental conditions
(Fig.~\ref{fig:paradigm}). The two conditions differed in how often we
cued participants to change the focus of their attention. All code and
documentation pertaining to our experiment and analyses, along with the
experimental stimuli and data, may be downloaded from
\url{http://www.github.com/ContextLab/attention-memory-task}.

@@

Participants viewed photographs of faces, places, and composite images each
comprising an equal blend of one face image and one place image. The pool of
360 face images included photographs of adult human male and female faces
selected from the FERET database~\citep{PhilEtal98}. The pool of 360 place
images included photographs of indoor and outdoor places selected from the SUN
database~\citep{XiaoEtal10}. The images we used from both databases came from a
stimulus subset that was manually curated by Megan deBettencourt (personal
communication). All images were resized to $256 \times 256$ pixels, converted
to greyscale, and processed so that every image was matched for mean contrast
and intensity. We selected 20 face images and 20 place images from the stimulus
pool to use in the instructional and practice phases of the experiment
(Fig.~\ref{fig:paradigm}A--D).

@@

We used the eye-tracking data to home in specifically on behavioral effects
related to \textit{covert} attention, as opposed to overt looking effects.
Specifically, we excluded from further analysis any images from trials where
participants shifted their gaze (for any non-zero amount of time) to any part
of the attended composite image during a presentation trial
(see Figs.~\gazeLocations,~and \excludedTrials).

@@

We also wondered whether the ways participants attended to or remembered the
images might depend on image-specific properties like the images' categories.
We repeated the analysis displayed in Figure~\ref{fig:familiarity} separately
for face and place images (Fig.~\ratingsByCategory). The same general patterns
reported above also held for each stimulus category individually. For example,
fully attended face and place images were both rated as more familiar than the
category-matched images from the unattended location (sustained: $t$s$(29) \geq
3.366, p$s$ \leq 0.002$; variable: $t$s$(22) \geq 4.773, p$s$ \leq 0.001$),
attended-location images from the unattended category (sustained: $t$s$(29)
\geq 5.886, p$s$ \leq 0.001$; variable: $t$s$(29) \geq 4.277, p$s$ \leq
0.001$), images from the unattended category and location (sustained: $t$s$(29)
\geq 6.628, p$s$ \leq 0.001$; variable: $t$s$(29) \geq 5.624, p$s$ \leq
0.001$), and novel images (sustained: $t$s$(29) \geq 5.987, p$s$ \leq 0.001$;
variable: $t$s$(29) \geq 5.132, p$s$ \leq 0.001$). Taken together, the above
results suggest that what we remember is guided in part by what we attended to,
even after accounting for where we look or what specifically we are looking at.

@@

Given the above results suggesting potential differences in the timescales of
category-based and location-based attention, we carried out two additional
exploratory analyses aimed at identifying other timing effects. First, we
wondered whether participants' familiarity ratings might show serial position
effects analogous to those reported in classic recognition memory
studies~\citep[e.g.,][]{Neat93b, McElDosh89, WickNorm66}. For each participant,
for each composite image pair (presented in each trial during the study phase
of the experiment), we labeled each composite's face and place image component
according to whether it matched the cued category and/or location. We discarded
any face or place images that did not appear in the participants' memory phase.
We tagged the remaining (probed) images with the familiarity ratings that
participants would later give the images during the memory phase and plotted
these ratings against the images' presentation positions
(Figs.~\ref{fig:timecourses}A and B). Across both experimental conditions and
across all serial positions, we generally found that the average ordering of
familiarity ratings by attention level (Fig.~\ref{fig:familiarity}) were
preserved. This suggests that the encoding-related affects of attention on
subsequent recognition memory are relatively stable over time (e.g., we did not
observe clear primacy or recency effects during the study phase of the
experiment). Second, we carried out an analogous analysis to identify potential
serial position effects of recall order. For each probed item a participant
rated during the memory phase, we assigned the image a label according to
whether the participant's attention cue (at the time the image was presented as
part of its composite pair) matched the image's category and/or location, or
whether the image was novel (i.e., not presented during the study phase).
Again, we found that (across both experimental conditions and all probe
positions) in general the average ordering of familiarity ratings by attention
level (Fig.~\ref{fig:familiarity}) were preserved. This suggests that the
retrieval-related affects of attention on subsequent recognition memory are
relatively stable over time (e.g., we did not observe clear primacy or recency
effects during the memory phase of the experiment).

Our finding that location-attended items appear to receive a familiarity boost
in the variable attention condition but not the sustained attention condition
is consistent with two possible interpretations. One possibility is that
focusing attention requires just a brief trigger (in this case, an attention
cue), but different forms of attention (in this case, category-based attention
versus location-based attention) require different amounts of time to ``ramp
up'' to full efficacy such that they begin to affect memory encoding. For
example, if category-based attention ramps up more slowly than location-based
attention, this might explain why the relative ordering of category-matched
versus location-matched unattended images changes between the sustained
attention and variable attention conditions (orange and yellow bars in
Fig.~\ref{fig:familiarity}). A second possibility is that each attention cue
provides a ``boost'' to memory encoding for the relevant aspects of one's
experience (e.g., image categories, spatial locations), but the size of the
boost varies across different forms of attention. If so, the \textit{number} of
successive attention cues one receives should predict how effectively the
attended and partially attended images are encoded. We developed a sequence
length analysis to distinguish between these possibilities. For each probed
(target) image that had been presented as a composite image pair, we computed
the number of matching attention cues the participant had received by the time
the given images were presented (up to and including the image's composite
pair). We computed these sequence lengths by defining ``matching'' cues in
three ways: (a) a match means the cues are for the same category \textit{and}
location, (b) a match means that the cues are for the same category, and (c) a
match means that the cues are for the same location. This yielded, for each
target image, an associated count of how many matching attention cues the
participant had received up to and including that image's presentation. As
shown in Figure~\sequenceLength, we used linear regressions of familiarity
ratings on sequence lengths to identify potential sequence length effects. For
both experimental conditions, all attention levels, and for all three
approaches to defining matching cue sequence lengths, we found virtually no
reliable associations between cue sequence length and familiarity. This finding
is most parsimonious with the first possibility mentioned above-- i.e., that
attention may be guided by a brief trigger, but that different forms of
attention may take differently long to affect memory encoding.

Finally, we wondered whether participants' familiarity ratings might be
influenced in part by response bias effects. For example, a participant who had
recently received an ``attend to face images'' cue might rate even a novel face
image as more ``familiar'' if they leveraged their memory of the attention cue
(as opposed to solely relying on their memories of the composite images they
studied) to help guide their familiarity ratings. Indeed, participants in the
sustained attention condition do show some response biases. For example,
participants tended to rate novel images as more familiar if they came from the
just-cued category (Fig.~\responseBias A; $t(29) = 4.371, p < 0.001$).
Responses biases are more difficult to evaluate in the variable attention
condition. For example, should cue recency be defined as the number studied
composite image pairs that came between the image whose familiarity the
participant is judging and the most recent same-category cue (i.e., distance to
the nearest same-category attention cue)? Or might response biases instead
arise when a given category is cued more often near the end of the just-studied
list? We assigned each probed image a label based on these two measures of
category-matched cue recency. We then used linear regressions of familiarity
ratings on these recency measures to identify potential response biases
(Fig.~\responseBias B). Altogether we found no evidence that participants in
the variable attention condition tended to rate images as more familiar solely
due to how recently they had received a same-category attention cue.

@@

  \caption{\textbf{How do covert location-based and category-based attention
  affect memory encoding?}\textbf{A. Hypothesized time courses of the impact of
  location-based and category-based attention on memory encoding.} Shifting the
  focus of location-based attention increases memory encoding at the attended
  locations (blue curve). This increase can be observed after 5.5~s (the
  duration of one presentation from the variable attention condition). Shifting
  the focus of category-based attention suppresses memory encoding for the
  unattended category. However, this suppression effect occurs relatively
  slowly (longer than the duration of a single image presentation in the
  experiment). \textbf{B. Location-based attention.} Focusing covert attention
  on one \textit{location} enhances encoding of stimuli at the attended
  location (red and yellow bars), regardless of stimulus category. (This panel
  is based on the variable attention results presented in
  Fig.~\ref{fig:familiarity}.) \textbf{C. Category-based attention.} Focusing
  covert attention on one \textit{category} suppresses encoding of stimuli from
  the unattended category (yellow and green bars), regardless of spatial
  location. (This panel is based on the sustained attention results presented
  in Fig.~\ref{fig:familiarity}.)}

@@

We also found that partially attended images (e.g., from the attended category
but unattended location, or at the attended location but unattended category)
were rated as more familiar than novel images. However, these encoding benefits
differed across the two experimental conditions. In the sustained attention
condition, attended-category images from the unattended location were rated as
more familiar than novel images, but attended-location images from the
unattended category were not. Participants in the variable attention condition
showed the opposite pattern. The variable attention participants rated
attended-\textit{location} images from the unattended category as more familiar
than novel images, but they rated attended-category images from the unattended
location similarly to novel images. Because the primary difference between the
sustained and variable attention conditions was the duration of participants'
focus of attention, our analyses of partially attended images suggest that the
effects of different aspects of attention on memory may unfold over different
timescales (Fig.~\ref{fig:discussion}). Specifically, location-based attention
appears to affect memory encoding relatively quickly, which would explain why
attended-location images from both the attended \textit{and} unattended
category receive a memory encoding benefit in the variable attention condition.
In contrast, category-based attention appears to affect memory encoding more
slowly, and it appears to operate in a ``suppressive'' manner (i.e.,
suppressing encoding of the unattended category, as opposed to enhancing
encoding of the attended category). This would explain why unattended-category
images at the attended location are rated as \textit{less} familiar in the
sustained attention condition.

The notion that location-based attention operates at a faster timescale than
category-based attention is supported by prior work on the deployment of visual
attention~\citep{SotoBlan04, StopEtal07}. Our findings that location-based
attention enhances the processing of attended stimuli whereas category-based
attention suppresses the processing of unattended stimuli is also consistent
with prior work on location-based attention~\citep[e.g.,][]{IttiKoch01} and
category-based attention~\citep[e.g.,][]{MoheEtal14}. Our finding that people
better remember attended stimuli also follows prior work on interactions
between attention and memory~\citep{PallWagn02, ChunTurk07, AlyTurk16,
AlyTurk17, WittEtal18, MorrEtal14, BaleEtal21}. Whereas much of this prior work
focused on elucidating the neural basis of these interactions, our work extends
these prior studies by elucidating the specific and separable behavioral
impacts of location-based attention (enhancement with a fast onset) and
category-based attention (inhibition with a slow onset) and on subsequent
recognition memory. Both of these effects persisted throughout the 2~min memory
phases of both conditions. Therefore future work is needed to elucidate the
longevity of these effects beyond 2 minutes.

@@

JRM and KZ developed the concept for this study. Experiment code was written by
KZ and ARM, and testing and data collection were conducted by MRL and KZ. KZ,
MRL, ARM, and JRM analyzed the data. JRM supervised the project. All authors
contributed to writing and editing the manuscript.

>>>>

Our ongoing subjective experiences, and our memories of those experiences, are
shaped by our prior experiences, goals, and situational understanding. These
factors influence how we allocate our attentional resources over different
aspects of our ongoing experiences. These attentional shifts may happen overtly
(e.g., when we change where we are looking) or covertly (e.g., without any
external physical manifestation). Additionally, we may attend to what is
happening at a specific spatial location (e.g., because we think something
important is happening there) or we may attend to particular features
irrespective of their locations (e.g., when we search for a friend's face in a
crowd versus a desired item in a grocery store). We ran a covert attention
experiment with two conditions that differed in how long they asked
participants to maintain the focus of the categories and locations they were
attending. Later, the participants performed a recognition memory task for
attended, unattended, and novel stimuli. Participants were able to shift the
location of their covert attentional focus more rapidly than they were able to
shift their focus of covert attention to stimulus categories, and the effects
of location-based attention on memory were longer-lasting than the effects of
category-based attention.

@@

We ran a total of 53 participants across two experimental conditions
(Fig.~\ref{fig:paradigm}). The two conditions differed in how often we cued
participants to change the focus of their attention. All code and documentation
pertaining to our experiment and analyses, along with the experimental stimuli
and data, may be downloaded from
\url{http://www.github.com/ContextLab/attention-memory-task}.

@@

Participants viewed photographs of faces, places, and composite images each
comprising an equal blend of one face image and one place image. The pool of
360 face images included photographs of adult human male and female faces
selected from the FERET database~\citep{PhilEtal98}. The pool of 360 place
images included photographs of indoor and outdoor scenes selected from the SUN
database~\citep{XiaoEtal10}. The images we used from both databases came from a
stimulus subset that was manually curated by Megan deBettencourt (personal
communication). All images were resized to $256 \times 256$ pixels, converted
to greyscale, and processed so that every image was matched for mean contrast
and intensity. We selected 20 face images and 20 place images from the stimulus
pool to use in the instructional and practice phases of the experiment
(Fig.~\ref{fig:paradigm}A--D).

@@

We used the eye-tracking data to home in on behavioral effects related to
\textit{covert} attention, as opposed to overt looking effects. Specifically,
we excluded from further analysis any images from trials where participants
shifted their gaze (for any non-zero amount of time) to any part of the
attended composite image during a presentation trial (see
Figs.~\gazeLocations,~and \excludedTrials).

@@

We also wondered whether the ways participants attended to or remembered the
images might depend on image-specific properties like the images' categories.
We repeated the analysis displayed in Figure~\ref{fig:familiarity} separately
for face and place images (Fig.~\ratingsByCategory). The same general patterns
reported above also held for each stimulus category individually. For example,
fully attended face and place images were both rated as more familiar than the
category-matched images from the unattended location (sustained: $t$s$(29) \geq
3.366,~p$s~$\leq 0.002$; variable: $t$s$(22) \geq 4.773,~p$s~$\leq 0.001$),
attended-location images from the unattended category (sustained: $t$s$(29)
\geq 5.886,~p$s~$\leq 0.001$; variable: $t$s$(29) \geq 4.277,~p$s~$\leq
0.001$), images from the unattended category and location (sustained: $t$s$(29)
\geq 6.628,~p$s~$\leq 0.001$; variable: $t$s$(29) \geq 5.624,~p$s~$\leq
0.001$), and novel images (sustained: $t$s$(29) \geq 5.987,~p$s~$\leq 0.001$;
variable: $t$s$(29) \geq 5.132,~p$s~$\leq 0.001$). Taken together, the above
results suggest that what we remember is guided in part by what we attended to,
even after accounting for where we look or what specifically we are looking at.

@@

Given the above results suggesting potential differences in the timescales of
category-based and location-based attention, we carried out two additional
exploratory analyses aimed at identifying other timing effects. First, we
wondered whether participants' familiarity ratings might show serial position
effects analogous to those reported in classic recognition memory
studies~\citep[e.g.,][]{Neat93b, McElDosh89, WickNorm66}. For each participant,
for each composite image pair (presented in each trial during the study phase
of the experiment), we labeled each composite's face and place image component
according to whether it matched the cued category and/or location. We discarded
any face or place images that did not appear in the participants' memory
phases. We tagged the remaining (probed) images with the familiarity ratings
that participants would later give the images during the memory phase and
plotted these ratings against the images' presentation positions
(Figs.~\ref{fig:timecourses}A and B). Across both experimental conditions and
across all serial positions, we generally found that the average ordering of
familiarity ratings by attention level (Fig.~\ref{fig:familiarity}) were
preserved. This suggests that the encoding-related affects of attention on
subsequent recognition memory are relatively stable over time (e.g., we did not
observe clear primacy or recency effects during the study phase of the
experiment). Second, we carried out an analogous analysis to identify potential
serial position effects of recall order. For each probed item a participant
rated during the memory phase, we assigned the image a label according to
whether the participant's attention cue (at the time the image was presented as
part of its composite pair) matched the image's category and/or location, or
whether the image was novel (i.e., not presented during the study phase).
Again, we found that (across both experimental conditions and all probe
positions) in general the average ordering of familiarity ratings by attention
level (Fig.~\ref{fig:familiarity}) were preserved. This suggests that the
retrieval-related affects of attention on subsequent recognition memory are
relatively stable over time (e.g., we did not observe clear primacy or recency
effects during the memory phase of the experiment).

Our finding that location-attended items appear to receive a familiarity boost
in the variable attention condition but not the sustained attention condition
is consistent with two possible interpretations. One possibility is that
focusing attention requires just a brief trigger (in this case, an attention
cue), but different forms of attention (in this case, category-based attention
versus location-based attention) require different amounts of time to ``ramp
up'' to full efficacy such that they begin to affect memory encoding. For
example, if category-based attention ramps up more slowly than location-based
attention, this might explain why the relative ordering of category-matched
versus location-matched unattended images changes between the sustained
attention and variable attention conditions (orange and yellow bars in
Fig.~\ref{fig:familiarity}). A second possibility is that each attention cue
provides a ``boost'' to memory encoding for the relevant aspects of one's
experience (e.g., image categories, spatial locations), but the size of the
boost varies across different forms of attention. If so, the \textit{number} of
successive attention cues one receives should predict how effectively the
attended and partially attended images are encoded. We developed a sequence
length analysis to distinguish between these possibilities. For each probed
(target) image that had been presented as a composite image pair, we computed
the number of matching attention cues the participant had received by the time
the given images were presented (up to and including the image's composite
pair). We computed these sequence lengths by defining ``matching'' cues in
three ways: (a) a match means the cues are for the same category \textit{and}
location, (b) a match means that the cues are for the same category, and (c) a
match means that the cues are for the same location. This yielded, for each
target image, an associated count of how many successive matching attention
cues the participant had received up to and including that image's
presentation. As shown in Figure~\sequenceLength, we used linear regressions of
familiarity ratings on sequence lengths to identify potential sequence length
effects. For both experimental conditions, all attention levels, and for all
three approaches to defining matching cue sequence lengths, we found virtually
no reliable associations between cue sequence length and familiarity. This
finding is most parsimonious with the first possibility mentioned above-- i.e.,
that attention may be guided by a brief trigger, but that different forms of
attention may take differently long to affect memory encoding.

Finally, we wondered whether participants' familiarity ratings might be
influenced in part by response bias effects. For example, a participant who had
recently received an ``attend to face images'' cue might rate even a novel face
image as more ``familiar'' if they leveraged their memory of the attention cue
(as opposed to solely relying on their memories of the composite images they
studied) to help guide their familiarity ratings. Indeed, participants in the
sustained attention condition do show some response biases. For example,
participants tended to rate novel images as more familiar if they came from the
just-cued category versus the unattended category (Fig.~\responseBias A; $t(29)
= 4.371, p < 0.001$). Responses biases are more difficult to evaluate in the
variable attention condition. For example, should cue recency be defined as the
number studied composite image pairs that came between the image whose
familiarity the participant is judging and the most recent same-category cue
(i.e., temporal distance to the nearest same-category attention cue)? Or might
response biases instead arise when a given category is cued more often near the
end of the just-studied list? We assigned each probed image a label based on
these two measures of category-matched cue recency. We then used linear
regressions of familiarity ratings on these recency measures to identify
potential response biases (Fig.~\responseBias B). Altogether we found no
evidence that participants in the variable attention condition tended to rate
images as more familiar solely due to how recently they had received a
same-category attention cue.

@@

  \caption{\textbf{How do covert location-based and category-based attention
  affect memory encoding?} \textbf{A.~Hypothesized time courses of the impact of
  location-based and category-based attention on memory encoding.} Shifting the
  focus of location-based attention increases memory encoding at the attended
  locations (blue curve). This increase can be observed after 5.5~s (the
  duration of one presentation from the variable attention condition). Shifting
  the focus of category-based attention suppresses memory encoding for the
  unattended category. However, this suppression effect occurs relatively
  slowly (longer than the duration of a single image presentation in the
  experiment). \textbf{B.~Location-based attention.} Focusing covert attention
  on one \textit{location} enhances encoding of stimuli at the attended
  location (red and yellow bars), regardless of stimulus category. (This panel
  is based on the variable attention results presented in
  Fig.~\ref{fig:familiarity}.) \textbf{C.~Category-based attention.} Focusing
  covert attention on one \textit{category} suppresses encoding of stimuli from
  the unattended category (yellow and green bars), regardless of spatial
  location. (This panel is based on the sustained attention results presented
  in Fig.~\ref{fig:familiarity}.)}

@@

We also found that partially attended images (e.g., from the attended category
but unattended location, or at the attended location but unattended category)
were rated as more familiar than novel images. However, these encoding benefits
differed across the two experimental conditions. In the sustained attention
condition, attended-category images from the unattended location were rated as
more familiar than novel images, but attended-location images from the
unattended category were not. Participants in the variable attention condition
showed the opposite pattern. In particular, variable attention participants
rated attended-\textit{location} images from the unattended category as more
familiar than novel images, but they rated attended-category images from the
unattended location similarly to novel images. Because the primary difference
between the sustained and variable attention conditions was the duration of
participants' focus of attention, our analyses of partially attended images
suggest that the effects of different aspects of attention on memory unfold
over different timescales (Fig.~\ref{fig:discussion}). Specifically,
location-based attention appears to affect memory encoding relatively quickly,
which would explain why attended-location images from both the attended
\textit{and} unattended category receive a memory encoding benefit in the
variable attention condition. In contrast, category-based attention appears to
affect memory encoding more slowly, and it appears to operate in a
``suppressive'' manner (i.e., suppressing encoding of the unattended category,
as opposed to enhancing encoding of the attended category). This would explain
why unattended-category images at the attended location are rated as
\textit{less} familiar in the sustained attention condition.

The notion that location-based attention operates at a faster timescale than
category-based attention is supported by prior work on the deployment of visual
attention~\citep{SotoBlan04, StopEtal07}. Our findings that location-based
attention enhances the processing of attended stimuli whereas category-based
attention suppresses the processing of unattended stimuli is also consistent
with prior work on location-based attention~\citep[e.g.,][]{IttiKoch01} and
category-based attention~\citep[e.g.,][]{MoheEtal14}. Our finding that people
better remember attended stimuli also follows prior work on interactions
between attention and memory~\citep{PallWagn02, ChunTurk07, AlyTurk16,
AlyTurk17, WittEtal18, MorrEtal14, BaleEtal21}. Whereas much of this prior work
focused on elucidating the neural basis of these interactions, our work extends
these prior studies by elucidating the specific and separable behavioral
impacts of location-based attention (enhancement with a fast onset) and
category-based attention (suppression with a slow onset) and on subsequent
recognition memory. Both of these effects persisted throughout the 2~min memory
phases of both conditions. However, future work is needed to elucidate the
longevity of these effects beyond 2 minutes.

@@

JRM and KZ developed the concept for this study. Experiment code was written by
ARM and KZ, and testing and data collection were conducted by KZ and MRL. All
authors contributed to analyzing the data. All authors contributed to writing
and editing the manuscript.  JRM supervised the project.

### 2023-03-27 paper/supplemental_materials.tex
minor wording tweaks, typo and formatting fixes

 \caption{\textbf{Gaze location histograms.} We divided the experimental
 display into 120 horizontal bins and 78 vertical bins, each comprising a
 roughly 0.5~cm square. The panels display the average proportions of time
 (throughout the duration of the experiment) participants spend looking at each
 location. The left panel displays data from participants in the sustained
 attention condition and the right panel displays data from participants in the
 variable attention condition. In both panels, the locations of the central
 fixation cross (red $+$) and composite images (red-outlined squares) are
 indicated.}

@@

\caption{\textbf{Proportions of excluded trials.} We excluded from our analyses
any images from trials where the participant's gaze touched on any part of the
attended composite image (attended-side red squares in
Fig.~\ref{fig:gaze-histograms}). \textbf{A--B. Proportions of excluded trials,
by run.} Across both experimental conditions (left: sustained attention; right:
variable attention), the bars display the average proportions of presentations
from each run where participants looked at any part of the attended composite
image, for any non-zero duration. Error bars denote across-participant
bootstrap-estimated 95\% confidence intervals. \textbf{C--D. Proportions of
excluded trials, by participant.} Across both experimental conditions (left:
sustained attention; right: variable attention), the bars display the average
proportions of presentations across different runs where the given
participants looked at any part of the attended composite image, for any
non-zero duration. Error bars denote across-run bootstrap-estimated 95\%
confidence intervals.}

@@

\caption{\textbf{Effects of attention cue sequence length on familiarity.} 
\textbf{A. Sequence effects (sustained attention condition).} Each heat map
displays the proportions of images that participants rated at each familiarity
rating (rows), broken down by the number of successive matching attention cues
the participant received at the time the given image was presented as part of a
composite pair (up to and including the image's composite pair; columns). The
trend lines (with bootstrap-estimated 95\% confidence intervals) show
regressions fit to the corresponding heatmap's data ($x$: sequence length; $y$:
familiarity rating). Heat maps are organized into rows based on how the sequence
lengths were computed (top row: successive attention cues had the same category
and location; middle row: successive attention cues had the same category;
bottom row: successive attention cues had the same location). Columns (and
colors) denote attention levels (as in Figs.~\familiarity~and
\ref{fig:ratings-by-category}). \textbf{B. Sequence effects (variable attention
condition).} This Panel is in the same format as Panel A, but displays results
from the variable attention condition.}

@@

\caption{\textbf{Response bias analyses.} \textbf{A. Sustained attention
condition.} The bars display the mean familiarity ratings participants gave to
novel memory probes (lures) that matched (``True'') versus did not match
(``False'') the prior run's attention-cued category. Error bars denote
across-participant bootstrap-estimated 95\% confidence intervals. \textbf{B.
Variable attention condition.} Each heat map displays the proportions of images
that participants rated at each familiarity rating (rows), broken down by
metrics related to recent same-category attention cues participants received on
the prior run (columns). The trend lines (with bootstrap-estimated 95\%
confidence intervals) show regressions fit to the corresponding heatmap's data
($x$: metrics related to same-category cues; $y$: familiarity rating). The top
panels's columns reflect the temporal distances to the most recent
same-category attention cues, measured in numbers of presented items (where 1
denotes the final item on the most recent run). The bottom panel's columns
denote the recency-weighted average numbers of same-category cues presented on
the most recent run. Each column of matrices corresponds to a different
attention level, as in Figures~\familiarity~and \ref{fig:ratings-by-category}.
\textbf{C. Weights by recency.} The panel displays the weights used to compute
recency-weighted numbers of same-category cues on the most recent run. The
$x$-axis displays the presentation positions of each attention cue, and the
$y$-axis displays the weight assigned to same-category cues at the given
presentation position. The weights are first computed using $w = \max\left[1 -
\exp\{-\frac{x}{\tau}\} , \epsilon \right]$, where $\tau = 2$, $\epsilon =
0.05$, and $x$ is the cue's presentation position. To obtain the $y$-values
shown in the Panel (and used in Panel B), the weights for $x \in \left[0, 1, 2,
..., 9\right]$ are normalized to sum to 1.}

>>>>

 \caption{\textbf{Gaze location histograms.} We divided the experimental
 display into 120 horizontal bins and 78 vertical bins, each comprising a
 roughly 0.5~cm square. The panels display the average proportions of time
 (throughout the duration of the experiment) participants spent looking at each
 location. The left panel displays data from participants in the sustained
 attention condition and the right panel displays data from participants in the
 variable attention condition. In both panels, the locations of the central
 fixation cross (red $+$) and composite images (red-outlined squares) are
 indicated.}

@@

\caption{\textbf{Proportions of excluded trials.} We excluded from our analyses
any images from trials where the participant's gaze touched on any part of the
attended composite image (attended-side red squares in
Fig.~\ref{fig:gaze-histograms}). \textbf{A--B. Proportions of excluded trials,
by trial block.} Across both experimental conditions (left: sustained
attention; right: variable attention), the bars display the average proportions
of presentations from each trial block where participants looked at any part of the
attended composite image, for any non-zero duration. Error bars denote
across-participant bootstrap-estimated 95\% confidence intervals. \textbf{C--D.
Proportions of excluded trials, by participant.} Across both experimental
conditions (left: sustained attention; right: variable attention), the bars
display the average proportions of presentations across different runs where
the given participants looked at any part of the attended composite image, for
any non-zero duration. Error bars denote across-block bootstrap-estimated 95\%
confidence intervals.}

  \label{fig:gaze-intersections}
\end{figure*}

@@

\caption{\textbf{Effects of attention cue sequence length on familiarity.} 
\textbf{A. Sustained attention condition.} Each heat map
displays the proportions of images that participants rated at each familiarity
rating (rows), broken down by the number of successive matching attention cues
the participant received at the time the given image was presented as part of a
composite pair (up to and including the image's composite pair; columns). The
trend lines (with bootstrap-estimated 95\% confidence intervals) show
regressions fit to the corresponding heatmap's data ($x$: sequence length; $y$:
familiarity rating). Heat maps are organized into rows based on how the sequence
lengths were computed (top row: successive attention cues had the same category
and location; middle row: successive attention cues had the same category;
bottom row: successive attention cues had the same location). Columns (and
colors) denote attention levels (as in Figs.~\familiarity~and
\ref{fig:ratings-by-category}). \textbf{B. Variable attention
condition.} This panel is in the same format as Panel A, but displays results
from the variable attention condition.}

@@

\caption{\textbf{Response bias analyses.} \textbf{A. Sustained attention
condition.} The bars display the mean familiarity ratings participants gave to
novel memory probes (lures) that matched (``True'') versus did not match
(``False'') the prior run's attention-cued category. Error bars denote
across-participant bootstrap-estimated 95\% confidence intervals. \textbf{B.
Variable attention condition.} Each heat map displays the proportions of images
that participants rated at each familiarity rating (rows), broken down by
metrics related to recent same-category attention cues participants received on
the prior run (columns). The trend lines (with bootstrap-estimated 95\%
confidence intervals) show regressions fit to the corresponding heatmap's data
($x$: metrics related to same-category cues; $y$: familiarity rating). The top
panels's columns reflect the temporal distances to the most recent
same-category attention cues, measured in numbers of presented items (where 1
denotes the final item on the most recent run). The bottom panel's columns
denote the recency-weighted average numbers of same-category cues presented on
the most recent run. Each column of matrices corresponds to a different
attention level, as in Figures~\familiarity~and \ref{fig:ratings-by-category}.
\textbf{C. Weights by recency.} The panel displays the weights used to compute
recency-weighted numbers of same-category cues on the most recent run. The
$x$-axis displays the presentation positions of each attention cue, and the
$y$-axis displays the weight assigned to same-category cues at the given
presentation position. The weights are first computed using $w = \max\left[1 -
\exp\{-\frac{x}{\tau}\} , \epsilon \right]$, where $\tau = 2$, $\epsilon =
0.05$, and $x$ is the cue's presentation position. To obtain the $y$-values
shown in the panel (and used in Panel B), the weights for $x \in \left[0, 1, 2,
..., 9\right]$ are normalized to sum to 1.}
